---
title: "Making Republic of Ireland CoM tables"
author: "Dave"
format: html
editor: visual
---

## Libraries

```{r libraries}
#| output: false
library(fs)
library(here)

library(vroom)
library(dplyr)
library(stringr)
library(purrr)
library(forcats)
library(tidyr)
library(snakecase)
```

## Read in data

### Read 'CoM List'

That is, the list which joins CoM to dioceses and nations.

```{r read_com_list}
com <- vroom(here("com-list.csv"), col_types = "cccc")

com <- filter(com, nation == "republic-of-ireland")

if(!interactive()) head(com, 3)
```

### Read CoM Stats

```{r define_stats}
stats <- c("age", "ethnicity",
           "ethnicity-harmonised", "general-health", 
           "household-composition", "religion")

names(stats) <- stats
```

```{r make_com_data_helper}
read_com_data <- function(x){
  
  census_file <-
    here("data", "com-stats",
         str_c("irish-com", x, "2016", sep = "_") |>
           str_c(".csv"))
  
  vroom(census_file, show_col_types = FALSE)
}
```

```{r read_com_data}
com_data <- map(stats, read_com_data)
```

### Read diocesan stats

```{r make_diocesan_stats_helper}
read_diocesan_data <- function(x){
  
  data_file <-
    here("data", "diocesan-stats",
         str_c("republic-of-ireland", "dioceses", x, "2016", sep = "_") |> 
           str_c(".csv"))
  
  vroom(data_file, show_col_types = FALSE)
}
```

```{r read_diocesan_stats}
diocesan_data <- map(stats, read_diocesan_data)

if(!interactive()){
  diocesan_data[[1]] |>
    head()
}
```

### Read national stats

```{r make_national_data_helper}
read_national_data <- function(x){
  
  data_file <- 
    here("data", "national-stats",
         str_c("republic-of-ireland", x, "sa", "2016", sep = "_") |> 
           str_c(".RDS"))
  
  readRDS(data_file)
}
```

```{r read_national_data}
national_data <- map(stats, read_national_data)

rm(stats)
```

## Prepare Joining Fields

```{r prepare_joining_fields}

diocesan_data <- map(diocesan_data,
                     \(x){
                       mutate(x, 
                              diocese =
                                str_to_lower(diocese) |>
                                str_squish() |> 
                                str_remove_all("[:punct:]") |> 
                                str_replace_all(" ", "-") |> 
                                str_remove_all("-and"))
                       })

national_data <-
  map(national_data,
      \(x){
        ungroup(x) |> 
          summarise(across(where(is.numeric), sum))
})

national_data <-
  map(national_data, \(x) mutate(x, nation = "republic-of-ireland"))

```

## Pivot tables longer and combine

### Hepler: `lenghten()`

```{r lengthen}

lengthen <- function(x, level, id_col){
  
  id_col <- enexpr(id_col)
  
  imap(x,
       \(table, stat_name){
         
         units <- ifelse(
           str_detect(stat_name, "household"), "households", "persons")
         
         pivot_longer(table, -!!id_col,
                      names_to = "category",
                      values_to = "count") |> 
           
           mutate(
             level := level,
             statistic := stat_name,
             units = units) |>
           relocate(level, !!id_col, statistic, category, units, count)
       })
}
```

### Pivot CoM data and combine

```{r pivot_com_data}

com_data_long <- lengthen(com_data, level = "centre-of-mission", com)
  
head(com_data_long[["household-composition"]])

com_data <- bind_rows(com_data_long)
```

### Pivot diocesan data and combine

```{r pivot_diocesan_data}

diocesan_data_long <- lengthen(diocesan_data, level = "diocese", diocese)

head(diocesan_data_long[["household-composition"]])

diocesan_data <- bind_rows(diocesan_data_long)
```

### Pivot national data and combine

```{r pivot_national_data}

national_data_long <- lengthen(national_data, level = "nation", nation)

head(national_data_long[["household-composition"]])

national_data <- bind_rows(national_data_long)
```

```{r clean_workspace_1}
rm(com_data_long, diocesan_data_long, national_data_long)
```

## Join CoM names to diocesan and national data

```{r join_com_names_to_data}

national_data <-
  right_join(national_data,
            select(com, com, nation),
            by = "nation") |> 
  select(-nation)

diocesan_data <- 
  right_join(diocesan_data,
            select(com, com, diocese),
            by = "diocese") |> 
  select(-diocese)

head(national_data)
head(diocesan_data)
head(com_data)
```

### Combine all data

```{r allData}
all_data <-
  bind_rows(com_data, diocesan_data, national_data) |> 
  relocate(com, level, statistic, category, units, count) |> 
  arrange(com, statistic, level, category)


glimpse(all_data)

all_data <- 
  left_join(com, all_data, by = "com") |> 
  mutate(nation_label = snakecase::to_title_case(nation)) |> 
  relocate(nation_label, .after = nation)

glimpse(all_data)
head(all_data)
```

## Save

```{r save}
save_dir <- dir_create(here("data", "comprehensive-com-data"))

saveRDS(all_data,
        file.path(save_dir, "republic-of-ireland_all-com-data_no-relative-counts.RDS"))

vroom_write(all_data, 
      file.path(save_dir, "republic-of-ireland_all-com-data_no-relative-counts.csv"),
      delim = ",")
```
